### 2022-08-19 prelim_results.Rmd
add ms figures file

---
title: "Seafood globalization preliminary results"
author: "Jessica Gephart"
output: 
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(countrycode)
library(circlize)
library(here)
library(ggpubr)
library(kableExtra)
library(DBI)
library(janitor)
library(viridis)
library(broom)
library(exploreARTIS)

```

```{r load_database_data}
# Load ARTIS data from SQL local database
con <- dbConnect(RPostgres::Postgres(),
                 dbname=Sys.getenv("POSTGRES_DB"),
                 host="localhost",
                 port="5432",
                 user=Sys.getenv("POSTGRES_USER"),
                 password=Sys.getenv("POSTGRES_PASSWORD"))

# Check that connection is established by checking which tables are present
dbListTables(con)

# Pull all ARTIS and production data
artis <- dbGetQuery(con, "SELECT * FROM snet")

artis <- artis %>%
  select(-record_id) %>%
  # Remove 2020 for now
  filter(year < 2020)

prod <- dbGetQuery(con, "SELECT * FROM production")
prod <- prod %>%
  select(-record_id) %>%
  # Remove 2020 for now
  filter(year < 2020)

# Close database connection
dbDisconnect(con)

rm(list = c("con"))
```

```{r set_directories}
# Set time series dates
date_nodash <- "20220614"
date_dash <- "2022-06-14"

# Set directories 
# datadir <-  paste("/Volumes/jgephart/ARTIS/Outputs/S_net/", date_nodash, "_ARTIS-timeseries", sep="")
model_inputs_dir <- "/Volumes/jgephart/ARTIS/Outputs/model_inputs_archive_20220718"
outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization"

# Set color palettes
# All colors from palette: #FFE6BE (light yellow), #F7AF75 (light orange), #F37542 (medium orange), 
# #E24027 (dark orange), #0F2D59 (navy), #1B708F (blue), #2E8D9A (teal), #86ADA7 (light blue)
method_colors <- c("#F7AF75", "#2E8D9A", "#86ADA7") # light orange, teal, light blue
source_colors <- c("#F37542", "#0F2D59", "#86ADA7") # medium orange, navy, light blue

```

```{r load_drive_data}
code_max_resolved_taxa <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/code_max_resolved_taxa.csv") %>%
              mutate(hs6 = as.character(hs6))

sciname_metadata <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/sciname_metadata.csv")

```

```{r join_data}
# Load cleaned FAO production data
prod <- prod %>% 
  select("year", "iso3c", "sciname", 
         "method" = "prod_method", "environment", 
         "production_t" = "live_weight_t") %>%
  group_by(year, iso3c, sciname, method, environment) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE)) %>%
  ungroup()

# Join production data to ARTIS
artis <- artis %>%
  filter(live_weight_t >= 0.1) %>%
# FIX IT: Determine if this should be deleted. Production of only species exports may not be relevant
  left_join(prod, by = c("year", "exporter_iso3c" = "iso3c", 
                           "sciname", "method", "environment")) %>%
  ungroup()

# Update artis data to lowest resolution by comparing code taxa_level and prod taxa_level
artis <- artis %>%
  # FIX IT: Change metadata to HSXX format and then remove this line
  mutate(hs_version = as.integer(parse_number(hs_version))) %>%
  left_join(code_max_resolved_taxa,
            by = c("hs_version", "hs6", "sciname")) %>%
  left_join(sciname_metadata, 
            by = c("sciname" = "SciName")) %>%
  # Leave any missing sciname_hs_modified as original sciname
  mutate(sciname_hs_modified = case_when(
    is.na(sciname_hs_modified) ~ sciname,
    TRUE ~ sciname_hs_modified
  ))

# Remove data frames after merged with ARTIS
rm(list = c("code_max_resolved_taxa", "sciname_metadata"))

```

# Trends in global seafood trade

```{r global_total_trend}
artis %>% 
  group_by(year) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000)) +
  geom_line() +
  labs(x = "", y = "Seafood exports (million t, live weight)") +
  theme_minimal()

artis_fmfo_ts <- artis %>% 
  mutate(food_or_fmfo = case_when(
    hs6 == "230120" ~ "fishmeal",
    hs6 != "230120" ~ "nonfishmeal"
  )) %>%
  group_by(year, food_or_fmfo) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE))

artis_fmfo_ts %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = food_or_fmfo)) +
  geom_area() +
  scale_fill_manual(values = method_colors) +
  labs(x = "", y = "Seafood exports (million t, live weight)", fill = "") +
  theme_minimal()

artis_fmfo_ts_summary <- artis_fmfo_ts %>%
  pivot_wider(names_from = food_or_fmfo, values_from = live_weight_t) %>%
  mutate(percent_fishmeal = 100*fishmeal/(fishmeal+nonfishmeal))

total_1996 <- round(sum(artis %>% filter(year == 1996) %>% pull(live_weight_t))/1000000, 1)
total_2019 <- round(sum(artis %>% filter(year == 2019) %>% pull(live_weight_t))/1000000, 1)

total_capture_1996 <- round(sum(artis %>% filter(year == 1996, method == "capture") %>% pull(live_weight_t))/1000000, 1)
total_capture_2019 <- round(sum(artis %>% filter(year == 2019, method == "capture") %>% pull(live_weight_t))/1000000, 1)

total_aquaculture_1996 <- round(sum(artis %>% filter(year == 1996, method == "aquaculture") %>% pull(live_weight_t))/1000000, 1)
total_aquaculture_2019 <- round(sum(artis %>% filter(year == 2019, method == "aquaculture") %>% pull(live_weight_t))/1000000, 1)

```

Global seafood trade increased from `r total_1996` million t (live weight) in 1996 to `r total_2019` million t (live weight) in 2019, representing a `r round(100*((total_2019-total_1996)/total_1996),0)`% increase over that period.

Global capture fishery trade increased from `r total_capture_1996` million t (live weight) in 1996 to `r total_capture_2019` million t (live weight) in 2019, representing a `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`% increase over that period. In 1996, capture fishery trade represented `r round(100*total_capture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_capture_2019/total_2019)`% of all seafood exports.

Global aquaculture fishery trade increased from `r total_aquaculture_1996` million t (live weight) in 1996 to `r total_aquaculture_2019` million t (live weight) in 2019, representing a `r round(100*((total_aquaculture_2019-total_aquaculture_1996)/total_aquaculture_1996),0)`% increase over that period. In 1996, aquaculture fishery trade represented `r round(100*total_aquaculture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_aquaculture_2019/total_2019)`% of all seafood exports.

```{r global_species_pattern}
n_species <- artis %>% 
  group_by(year, sciname_hs_modified) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  filter(live_weight_t > 1000) %>%
  group_by(year) %>% 
  tally()

n_species %>%
  ggplot(aes(x = year, y = n)) +
  geom_line() +
  labs(x = "", y = "Number of species/groups with >1000 t exported") +
  theme_minimal()

n_species_1996 <- sum(n_species %>% filter(year == 1996) %>% pull(n))
n_species_2019 <- sum(n_species %>% filter(year == 2019) %>% pull(n))

```

The number of species/species groups identified in exports increased from `r n_species_1996` m in 1996 to `r n_species_2019` in 2019, representing a `r round(100*((n_species_2019-n_species_1996)/n_species_1996),0)`% increase over that period. Increases in species/species groups exported can arise from either an increase in the species traded internationally or from an increase in the resolution of the trade and production data over time.

# Top traded species

```{r exports_by_species}
# Top species exported
top_species_1 <- artis %>% 
  filter(year %in% 1996:2000) %>%
  filter(str_count(sciname_hs_modified, pattern = " ") == 1) %>% 
  group_by(sciname_hs_modified, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(sciname_hs_modified) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(sciname_hs_modified = fct_reorder(as.factor(sciname_hs_modified), total)) %>%
  slice_max(n = 31, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = sciname_hs_modified, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 species \nexported (1996-2000)", fill = "Production method") +
  theme_minimal()

top_species_2 <- artis %>% 
  filter(year %in% 2015:2019) %>%
  filter(str_count(sciname_hs_modified, pattern = " ") == 1) %>% 
  group_by(sciname_hs_modified, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(sciname_hs_modified) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(sciname_hs_modified = fct_reorder(as.factor(sciname_hs_modified), total)) %>%
  slice_max(n = 31, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = sciname_hs_modified, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 species \nexported (2015-2019)", fill = "Production method") +
  theme_minimal()

# Top species groups exported
top_species_groups_1 <- artis %>% 
  filter(year %in% 1996:2000) %>%
  filter(str_count(sciname_hs_modified, pattern = " ") == 0) %>% 
  group_by(sciname_hs_modified, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(sciname_hs_modified) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(sciname_hs_modified = fct_reorder(as.factor(sciname_hs_modified), total)) %>%
  slice_max(n = 31, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = sciname_hs_modified, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 groups \nexported (1996-2000)", fill = "Production method") +
  theme_minimal()

top_species_groups_2 <- artis %>% 
  filter(year %in% 2015:2019) %>%
  filter(str_count(sciname_hs_modified, pattern = " ") == 0) %>% 
  group_by(sciname_hs_modified, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(sciname_hs_modified) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(sciname_hs_modified = fct_reorder(as.factor(sciname_hs_modified), total)) %>%
  slice_max(n = 31, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = sciname_hs_modified, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 groups \nexported (2015-2019)", fill = "Production method") +
  theme_minimal()

ggarrange(top_species_groups_1, top_species_groups_2, common.legend = TRUE)
ggarrange(top_species_1, top_species_2, common.legend = TRUE)

```

# How does capture versus aquaculture trade globalization compare in terms of quantity and connectedness and how has this changed over time?

```{r}
# Stacked lines, colored by method
artis %>% 
  group_by(year, method) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = method)) +
  geom_area() +
  scale_fill_manual(values = method_colors) +
  labs(x = "", y = "Seafood exports (million t, live weight)", "Production method") +
  theme_minimal()

# Line color by method
artis %>% 
  group_by(year, method) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, color = method)) +
  geom_line() +
  scale_color_manual(values = method_colors) +
  labs(x = "", y = "Seafood exports (million t, live weight)", fill = "Production method") +
  theme_minimal()

# Summary stats
## Percent of exports by source
kable(artis %>% 
        group_by(year, method) %>%
        summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
        mutate(percent = round(100*live_weight_t/sum(live_weight_t), 1)) %>%
        select(year, method, percent) %>% 
        pivot_wider(names_from = method, values_from = percent),
      caption = "Percent of exports by production source")

## Number of exporters by source
kable(artis %>% 
        filter(live_weight_t>10) %>%
        select(year, exporter_iso3c, method) %>%
        distinct() %>%
        group_by(year, method) %>%
        tally() %>%
        select(year, method, n) %>% 
        pivot_wider(names_from = method, values_from = n),
      caption = "Number of exporters by production source")

## Number of importers by source
kable(artis %>% 
        select(year, importer_iso3c, method) %>%
        distinct() %>%
        group_by(year, method) %>%
        tally() %>%
        select(year, method, n) %>% 
        pivot_wider(names_from = method, values_from = n),
      caption = "Number of importers by production source")

# Average out degree by source
kable(artis %>% 
        select(year, exporter_iso3c, importer_iso3c, method) %>%
        distinct() %>%
        group_by(year, exporter_iso3c, method) %>%
        tally() %>%
        group_by(year, method) %>%
        summarise(n = round(mean(n), 1)) %>%
        select(year, method, n) %>% 
        pivot_wider(names_from = method, values_from = n),
      caption = "Average out degree by production source")

# Out degree distribution by year
artis %>% 
  select(year, exporter_iso3c, importer_iso3c, method) %>%
  distinct() %>%
  group_by(year, exporter_iso3c, method) %>%
  tally() %>%
  ggplot(aes(n, fill = method)) +
  geom_density(alpha = .5) +
  scale_fill_manual(values = method_colors) +
  facet_wrap(~year) +
  labs(title = "Out degree distribution by method", fill = "Production method")

# Average in degree by source
kable(artis %>% 
        select(year, exporter_iso3c, importer_iso3c, method) %>%
        distinct() %>%
        group_by(year, importer_iso3c, method) %>%
        tally() %>%
        group_by(year, method) %>%
        summarise(n = round(mean(n), 1)) %>%
        select(year, method, n) %>% 
        pivot_wider(names_from = method, values_from = n),
      caption = "Average in degree by production source")

# In degree distribution by year
artis %>% 
  select(year, exporter_iso3c, importer_iso3c, method) %>%
  distinct() %>%
  group_by(year, importer_iso3c, method) %>%
  tally() %>%
  ggplot(aes(n, fill = method)) +
  geom_density(alpha = .5) +
  scale_fill_manual(values = method_colors) +
  facet_wrap(~year) +
  labs(title = "Out degree distribution by method", fill = "Production method")

```

# How have regional trade patterns changed over time?

```{r}
# Top exporters, averaged over beginning and end time periods
top_exp_1 <- artis %>% 
  filter(year %in% 1996:2000) %>%
  group_by(exporter_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(exporter_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 seafood \nexporters (1996-2000)", fill = "Production method") +
  theme_minimal()

top_exp_2 <- artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(exporter_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(exporter_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 seafood \nexporters (2015-2019)", fill = "Production method") +
  theme_minimal()

ggarrange(top_exp_1, top_exp_2, common.legend = TRUE)

# Top importers, averaged over beginning and end time periods
top_imp_1 <- artis %>% 
  filter(year %in% 1996:2000) %>%
  group_by(importer_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(importer_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = fct_reorder(as.factor(importer_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood imports (million t, live weight)", y = "",
       title = "Top 20 seafood \nimporters (1996-2000)", fill = "Production method") +
  theme_minimal()

top_imp_2 <- artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(importer_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(importer_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = fct_reorder(as.factor(importer_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood imports (million t, live weight)", y = "",
       title = "Top 20 seafood \nimporters (2015-2019)", fill = "Production method") +
  theme_minimal()

ggarrange(top_imp_1, top_imp_2, common.legend = TRUE)

```

```{r capture_aqua_sankey, eval = FALSE}
artis %>% 
  filter(method == "capture", year %in% 2015:2019) %>% 
  plot_sankey(prop_flow_cutoff = 0.03)

artis %>% 
  filter(method == "aquaculture", year %in% 2015:2019) %>% 
  plot_sankey(prop_flow_cutoff = 0.03)
```

```{r regional_bilateral_trends}
artis_region <- artis %>%
  mutate(importer_region = countrycode(importer_iso3c, origin = "iso3c", destination = "region"),
         exporter_region = countrycode(exporter_iso3c, origin = "iso3c", destination = "region")) %>%
  filter(!is.na(importer_region), !is.na(exporter_region)) %>%
  group_by(year, exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_importer = paste(exporter_region, " to ", importer_region, sep = ""))

region_slopes <- artis_region %>%
  split(.$exporter_importer) %>% 
  map(~lm(live_weight_t~year, data = .x)) %>% 
  map_df(broom::tidy, .id = 'exporter_importer') %>%
  filter(term == 'year') %>%
  select(exporter_importer, "slope" = "estimate")

artis_region <- artis_region %>%
  left_join(region_slopes, by = "exporter_importer")
```

```{r region_trend_lollipop, fig.height=7, fig.width=5}
region_slopes %>%
  mutate(exporter_importer = fct_reorder(exporter_importer, slope)) %>%
  ggplot(aes(x = slope/1000, y = exporter_importer, color = slope, fill = slope)) +
  geom_segment(aes(xend = 0, yend = exporter_importer)) +
  geom_point() +
  scale_color_gradient2(low = "#E24027", mid = "grey70", high = "#1B708F") +
  scale_fill_gradient2(low = "#E24027", mid = "grey70", high = "#1B708F") +
  labs(y = "", x = "Ave. annual change in export (1000 t live weight)") +
  theme_bw() +
  theme(legend.position = "blank")

```

```{r region_trend_matrix, fig.height=7, fig.width=7}
artis_region %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, color = slope/1000)) +
  geom_line() +
  scale_color_gradient2(low = "#E24027", mid = "grey70", high = "#1B708F") +
  labs(x = "", y = "Exports (million t)", color = "Ave. annual change in export \n(1000 t live weight)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        legend.position = "bottom") +
  facet_grid(rows = vars(exporter_region), cols = vars(importer_region), scales = "free",
             labeller = label_wrap_gen(width = 5))

```

```{r region_trend_heatmap}
region_slopes %>%
  separate(exporter_importer, into = c("exporter", "importer"), sep = " to ") %>% 
  ggplot(aes(x = exporter, y = importer, fill = slope/1000)) +
  geom_tile() +
  scale_fill_gradient2(low = "#E24027", mid = "grey90", high = "#1B708F") +
  labs(x = "Exporting region", y = "Importing region", fill = "Ave. annual change in export \n(1000 t live weight)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        legend.position = "bottom")
```

```{r summarize_region_slopes}
kable(region_slopes %>%
  separate(exporter_importer, into = c("exporter", "importer"), sep = " to ") %>% 
  mutate(slope_dir = case_when(
    slope > 0 ~ 1, 
    slope < 0 ~ 0
  )) %>% 
  group_by(exporter) %>%
  summarise(total_increases = sum(slope_dir)) %>%
  arrange(desc(total_increases)), col.names = c("Exporting region", "Number of increases"))

```

```{r econ_group_trends}
# LCD countries according to UN (https://www.un.org/development/desa/dpad/least-developed-country-category/ldcs-at-a-glance.html)
LCD_countries <- c("Afghanistan", "Angola", "Bangladesh", "Benin", "Bhutan", "Burkina Faso", "Burundi", "Cambodia", "Central African Republic", "Chad", "Comoros", "Democratic Republic of the Congo", "Djibouti", "Eritrea", "Ethiopia", "Gambia", "Guinea", "Guinea-Bissau", "Haiti", "Kiribati", "Lao People’s Dem. Republic", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mozambique", "Myanmar", "Nepal", "Niger", "Rwanda", "Sao Tome and Principe", "Senegal", "Sierra Leone", "Solomon Islands", "Somalia", "South Sudan", "Sudan", "Timor-Leste", "Togo", "Tuvalu", "Uganda", "United Republic of Tanzania", "Yemen", "Zambia")

LCD_countries <- countrycode(LCD_countries, origin = "country.name", destination = "iso3c")

artis_lcd <- artis %>%
  mutate(
    importer_region = case_when(
    (importer_iso3c %in% LCD_countries) ~ "LCD",
    !(importer_iso3c %in% LCD_countries) ~ "non-LCD"),
     exporter_region = case_when(
    (exporter_iso3c %in% LCD_countries) ~ "LCD",
    !(exporter_iso3c %in% LCD_countries) ~ "non-LCD")) %>%
  group_by(year, exporter_region, importer_region, method) %>%
  summarise(live_weight_t = sum(live_weight_t))

artis_lcd %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = method)) +
  geom_area() +
  scale_fill_manual(values = method_colors) +
  labs(x = "", y = "Exports (million t)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_grid(rows = vars(exporter_region), cols = vars(importer_region))

artis_lcd %>% 
  group_by(importer_region, year, method) %>%
  summarise(total_imports = sum(live_weight_t)) %>%
  rename("region" = "importer_region") %>%
  left_join(artis_lcd %>% 
              group_by(exporter_region, year, method) %>%
              summarise(total_exports = sum(live_weight_t)) %>%
              rename("region" = "exporter_region"),
            by = c("year", "region", "method")) %>%
  filter(region == "LCD") %>%
  mutate(net_export = total_exports - total_imports) %>%
  ggplot(aes(x = year, y = net_export, color = method)) +
  scale_color_manual(values = method_colors) +
  geom_line() +
  labs(x = "", y = "Net export (t, live weight)") +
  ylim(c(0, 3100000)) +
  theme_bw()
```

```{r top_bilateral_flows, fig.height=4, fig.width=7}
top_flows_1 <- artis %>%
  filter(year %in% 1996:2000) %>% 
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)/5) %>%
  arrange(desc(live_weight_t)) %>%
  ungroup() %>%
  slice_max(n = 25, order_by = live_weight_t) %>%
  mutate(exporter_to_importer = paste(
    countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name"), 
    "to",
    countrycode(importer_iso3c, origin = "iso3c", destination = "country.name"),
    sep = " ")) %>%
  mutate(exporter_to_importer = fct_reorder(exporter_to_importer, live_weight_t)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_to_importer)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 25 bilateral \nflows (1996-2000)", x = "Average trade (million t)", y = "") +
  theme_bw()

top_flows_2 <- artis %>%
  filter(year %in% 2015:2019) %>% 
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)/5) %>%
  arrange(desc(live_weight_t)) %>%
  ungroup() %>%
  slice_max(n = 25, order_by = live_weight_t) %>%
  mutate(exporter_to_importer = paste(
    countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name"), 
    "to",
    countrycode(importer_iso3c, origin = "iso3c", destination = "country.name"),
    sep = " ")) %>%
  mutate(exporter_to_importer = fct_reorder(exporter_to_importer, live_weight_t)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_to_importer)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 25 bilateral \nflows (2015-2019)", x = "Average trade (million t)", y = "") +
  theme_bw()

ggarrange(top_flows_1, top_flows_2, nrow = 1)

```

```{r country_bilateral_trends, eval=FALSE}

country_slopes <- artis %>%
  mutate(exporter_importer = paste(exporter_iso3c, " to ", importer_iso3c, sep = "")) %>%
  split(.$exporter_importer) %>% 
  map(~lm(live_weight_t~year, data = .x)) %>% 
  map_df(broom::tidy, .id = 'exporter_importer') %>%
  filter(term == 'year') %>%
  select(exporter_importer, "slope" = "estimate")

```

# Exports from domestic, imported and unknown sources

```{r exports_by_source}
# Stacked lines, colored by export source
artis %>% 
  group_by(year, dom_source) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = dom_source)) +
  geom_area() +
  scale_fill_manual(values = source_colors) +
  labs(x = "", y = "Seafood exports (million t, live weight)", fill = "Export source") +
  theme_minimal()

# Line color by export source
artis %>% 
  group_by(year, dom_source) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, color = dom_source)) +
  geom_line() +
  scale_color_manual(values = source_colors) +
  labs(x = "", y = "Seafood exports (million t, live weight)", fill = "Export source") +
  theme_minimal()

## Percent of exports by source
kable(artis %>% 
        group_by(year, dom_source) %>%
        summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
        mutate(percent = round(100*live_weight_t/sum(live_weight_t), 1)) %>%
        select(year, dom_source, percent) %>% 
        pivot_wider(names_from = dom_source, values_from = percent),
      caption = "Percent of exports by export source")

total_domestic_1996 <- round(sum(artis %>% filter(year == 1996, dom_source == "domestic export") %>% pull(live_weight_t))/1000000, 1)
total_domestic_2019 <- round(sum(artis %>% filter(year == 2019, dom_source == "domestic export") %>% pull(live_weight_t))/1000000, 1)

total_foreign_1996 <- round(sum(artis %>% filter(year == 1996, dom_source == "foreign export") %>% pull(live_weight_t))/1000000, 1)
total_foreign_2019 <- round(sum(artis %>% filter(year == 2019, dom_source == "foreign export") %>% pull(live_weight_t))/1000000, 1)

```

Global domestic exports increased from `r total_domestic_1996` million t (live weight) in 1996 to `r total_domestic_2019` million t (live weight) in 2019, representing a `r round(100*((total_domestic_2019-total_domestic_1996)/total_domestic_1996),0)`% increase over that period. In 1996, domestic exports represented `r round(100*total_domestic_1996/total_1996)`% of all  exports and by 2019 domestic exports represented `r round(100*total_domestic_2019/total_2019)`% of all seafood exports.

Global foreign (i.e., imported and re-exported) exports increased from `r total_foreign_1996` million t (live weight) in 1996 to `r total_foreign_2019` million t (live weight) in 2019, representing a `r round(100*((total_foreign_2019-total_foreign_1996)/total_foreign_1996),0)`% increase over that period. In 1996, foreign exports represented `r round(100*total_foreign_1996/total_1996)`% of all  exports and by 2019 foreign exports represented `r round(100*total_foreign_2019/total_2019)`% of all seafood exports.

# How has foreign dependence evolved over time?

```{r exports_percent_by_source}
annual_total_prod <- prod %>%
  select(year, iso3c, sciname, production_t) %>%
  distinct() %>%
  group_by(year, sciname) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE)) %>%
  ungroup(sciname) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE))

# Stacked lines, colored by export source
artis %>% 
  group_by(year, dom_source) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  left_join(annual_total_prod, by = "year") %>%
  mutate(percent_export = 100*live_weight_t/production_t) %>%
  ggplot(aes(x = year, y = percent_export, fill = dom_source)) +
  geom_area() +
  scale_fill_manual(values = source_colors) +
  labs(x = "", y = "Seafood exports (million t, live weight)", fill = "Export source") +
  theme_minimal()

# Line color by export source
artis %>% 
  group_by(year, dom_source) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  left_join(annual_total_prod, by = "year") %>%
  mutate(percent_export = 100*live_weight_t/production_t) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, color = dom_source)) +
  geom_line() +
  scale_color_manual(values = source_colors) +
  labs(x = "", y = "Seafood exports (million t, live weight)", fill = "Export source") +
  theme_minimal()

## Percent of exports by source
kable(artis %>% 
        group_by(year, dom_source) %>%
        summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
        ungroup() %>%
        left_join(annual_total_prod, by = "year") %>%
        mutate(percent = round(100*live_weight_t/production_t, 1)) %>%
        select(year, dom_source, percent) %>% 
        pivot_wider(names_from = dom_source, values_from = percent),
      caption = "Percent of exported production by export source")
```

```{r exports_percent_by_species, fig.width=5, fig.height=7}
prod_1 <- artis %>%
  filter(dom_source == "domestic export", 
         year %in% 1996:2000) %>%
  select(exporter_iso3c, sciname, production_t) %>%
  distinct() %>%
  group_by(sciname) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE))

percent_export_1 <- artis %>% 
  filter(dom_source == "domestic export", 
         year %in% 1996:2000) %>% 
  group_by(sciname, CommonName) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  left_join(prod_1, by = c("sciname")) %>%
  mutate(percent_exported = 100*live_weight_t/production_t) %>%
  ungroup() %>%
  slice_max(n = 25, order_by = percent_exported) %>%
  mutate(CommonName = fct_reorder(as.factor(CommonName), percent_exported)) %>%
    mutate(percent_exported = case_when(
    percent_exported > 100 ~ 100,
    TRUE ~ percent_exported
  )) %>%
  ggplot(aes(x = percent_exported, y = CommonName)) +
  geom_bar(stat="identity", position = "stack") + 
  labs(x = "Percent of production exported as domestic export", y = "",
       title = "Top 25 species \n(1996-2000)") +
  theme_minimal()

prod_2 <- artis %>%
  filter(dom_source == "domestic export", 
         year %in% 2015:2019) %>%
  select(exporter_iso3c, sciname, production_t) %>%
  distinct() %>%
  group_by(sciname) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE))

percent_export_2 <- artis %>% 
  filter(dom_source == "domestic export", 
         year %in% 2015:2019) %>% 
  group_by(sciname, CommonName) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  left_join(prod_2, by = c("sciname")) %>%
  mutate(percent_exported = 100*live_weight_t/production_t) %>%
  mutate(percent_exported = case_when(
    percent_exported > 100 ~ 100,
    TRUE ~ percent_exported
  )) %>%
  ungroup() %>%
  slice_max(n = 25, order_by = percent_exported) %>%
  mutate(CommonName = fct_reorder(as.factor(CommonName), percent_exported)) %>%
  ggplot(aes(x = percent_exported, y = CommonName)) +
  geom_bar(stat="identity", position = "stack") + 
  labs(x = "Percent of production exported as domestic export", y = "",
       title = "Top 25 species \n(2015-2019)") +
  theme_minimal()

ggarrange(percent_export_1, percent_export_2, common.legend = TRUE)

```

```{r exports_percent_by_country}
prod_1 <- prod %>%
  filter(year %in% 1996:2000) %>%
  select(iso3c, sciname, production_t) %>%
  distinct() %>%
  group_by(iso3c) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE))

percent_export_1 <- artis %>% 
  filter(dom_source == "domestic export", 
         year %in% 1996:2000) %>% 
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  left_join(prod_1, by = c("exporter_iso3c" = "iso3c")) %>%
  mutate(percent_exported = 100*live_weight_t/production_t) %>%
  ungroup() %>%
  slice_max(n = 25, order_by = percent_exported) %>%
  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), percent_exported)) %>%
    mutate(percent_exported = case_when(
    percent_exported > 100 ~ 100,
    TRUE ~ percent_exported
  )) %>%
  mutate(country_name = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  ggplot(aes(x = percent_exported, y = exporter_iso3c)) +
  geom_bar(stat="identity", position = "stack") + 
  labs(x = "Percent of production exported \nas domestic export", y = "",
       title = "Top 25 countries \n(1996-2000)") +
  theme_minimal()

prod_2 <- prod %>%
  filter(year %in% 2015:2019) %>%
  select(iso3c, sciname, production_t) %>%
  distinct() %>%
  group_by(iso3c) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE))

percent_export_2 <- artis %>% 
  filter(dom_source == "domestic export", 
         year %in% 2015:2019) %>% 
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  left_join(prod_2, by = c("exporter_iso3c" = "iso3c")) %>%
  mutate(percent_exported = 100*live_weight_t/production_t) %>%
  ungroup() %>%
  slice_max(n = 25, order_by = percent_exported) %>%
  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), percent_exported)) %>%
    mutate(percent_exported = case_when(
    percent_exported > 100 ~ 100,
    TRUE ~ percent_exported
  )) %>%
  mutate(country_name = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  ggplot(aes(x = percent_exported, y = exporter_iso3c)) +
  geom_bar(stat="identity", position = "stack") + 
  labs(x = "Percent of production exported \nas domestic export", y = "",
       title = "Top 25 countries \n(2015-2019)") +
  theme_minimal()

ggarrange(percent_export_1, percent_export_2, common.legend = TRUE)

```

```{r calculate_supply}
exports <- artis %>%
  group_by(year, exporter_iso3c, dom_source, sciname, CommonName, method, environment) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  mutate(dom_source = str_replace(dom_source, " ", "_")) %>%
  pivot_wider(names_from = "dom_source", values_from = "live_weight_t")

imports <- artis %>%
  group_by(year, importer_iso3c, sciname, CommonName, method, environment) %>%
  summarise(imports_live_weight_t = sum(live_weight_t))

supply <- exports %>%
  full_join(imports, by = c("year", "exporter_iso3c" = "importer_iso3c", 
                            "sciname", "CommonName", "method", "environment")) %>%
  full_join(prod, by = c("year", "exporter_iso3c" = "iso3c", 
                          "sciname", "method", "environment"))

supply$foreign_export[is.na(supply$foreign_export)] <- 0
supply$domestic_export[is.na(supply$domestic_export)] <- 0
supply$error_export[is.na(supply$error_export)] <- 0
supply$imports_live_weight_t[is.na(supply$imports_live_weight_t)] <- 0
supply$production_t[is.na(supply$production_t)] <- 0

supply <- supply %>%
  mutate(supply = production_t + imports_live_weight_t - domestic_export - foreign_export - error_export,
         supply_no_error = production_t + imports_live_weight_t - domestic_export - foreign_export, 
         supply_domestic = production_t - domestic_export,
         supply_foreign = imports_live_weight_t - foreign_export) %>%
  rename("iso3c" = "exporter_iso3c") %>%
  ungroup() %>%
  mutate(supply_no_error = replace_na(supply_no_error, 0),
         supply_domestic = replace_na(supply_domestic, 0),
         supply_foreign = replace_na(supply_foreign, 0))

```

```{r supply_source_by_country}
select_countries <- c("CHN", "USA", "CHL", "GHA", "AUS", "NOR", "SWE", "MEX", "THA")

# Line graph of total supply for select countries
supply %>% 
  filter(iso3c %in% select_countries) %>%
  mutate(supply_no_error = case_when(supply_no_error < 0 ~0,
                                     TRUE ~ supply_no_error)) %>%
  group_by(iso3c, year) %>%
  summarise(supply_no_error = sum(supply_no_error)) %>%
  ggplot(aes(x = year, y = supply_no_error)) +
  geom_line() +
  labs(x = "", y = "Supply (t)") +
  facet_wrap(~iso3c) +
  theme_bw()

# Percent of supply by source for select countries
supply %>% 
  filter(iso3c %in% select_countries) %>%
  mutate(supply_no_error = case_when(supply_no_error < 0 ~ 0,
                                     TRUE ~ supply_no_error)) %>%
  group_by(iso3c, year) %>%
  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
  ggplot() +
  geom_line(aes(x = year, y = supply_domestic), color = "#1B708F") +
  geom_line(aes(x = year, y = supply_foreign), color = "#E24027") +
  labs(x = "", y = "Percent of supply (blue = domestic; red = foreign)") +
  facet_wrap(~iso3c) +
  theme_bw()
```

```{r supply_source_by_region}

supply %>% 
  mutate(region = countrycode(iso3c, origin = "iso3c", destination = "region")) %>%
  mutate(supply_no_error = case_when(supply_no_error < 0 ~ 0,
                                     TRUE ~ supply_no_error)) %>%
  group_by(region, year) %>%
  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
  filter(region != "NA") %>%
  ggplot() +
  geom_line(aes(x = year, y = supply_domestic), color = "#1B708F") +
  geom_line(aes(x = year, y = supply_foreign), color = "#E24027") +
  labs(x = "", y = "Percent of supply (blue = domestic; red = foreign)") +
  facet_wrap(~region) +
  theme_bw()

supply %>% 
  mutate(supply_no_error = case_when(supply_no_error < 0 ~ 0,
                                     TRUE ~ supply_no_error)) %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  #filter(sum(supply_no_error) > 100000) %>%
  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
  arrange(desc(supply_foreign))

```

```{r supply_source_by_species}
supply_species <- supply %>% 
  group_by(sciname, CommonName, year) %>%
  mutate(total_production_t = sum(production_t)) %>%
  summarise(percent_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            percent_foreign = 100*sum(supply_foreign)/sum(supply_no_error), 
            production_t = sum(production_t)) %>%
  arrange(desc(percent_foreign))

supply_species %>% 
  group_by(sciname, CommonName) %>% 
  mutate(total_production_t = sum(production_t)) %>%
  ungroup() %>%
  slice_max(n = 16*length(unique(supply_species$year)), order_by = total_production_t) %>%
  ggplot() +
  geom_line(aes(x = year, y = percent_domestic), color = "#1B708F") + 
  geom_line(aes(x = year, y = percent_foreign), color = "#E24027") +
  labs(x = "", y = "Percent of supply (blue = domestic; red = foreign)") +
  facet_wrap(~CommonName) +
  theme_bw()

```

# Trends in estimated error

```{r explore_country_error, eval=FALSE}
# Plot time series of species with largest volume error exports
supply %>%
  group_by(year, iso3c) %>%
  summarise(error_export = sum(error_export, na.rm = TRUE), 
            domestic_export = sum(domestic_export, na.rm = TRUE),
            foreign_export = sum(foreign_export, na.rm = TRUE),
            error_percent = 100*sum(error_export, na.rm = TRUE)/(sum(domestic_export, na.rm = TRUE) + 
                                                     sum(foreign_export, na.rm = TRUE) + 
                                                     sum(error_export, na.rm = TRUE))) %>%
  group_by(iso3c) %>%
  mutate(total_error = sum(error_export)) %>%
  ungroup() %>%
  slice_max(n = 12*length(unique(supply$year)), order_by = total_error) %>%
  ggplot(aes(x = year, y = error_percent)) +
  geom_line() +
  labs(x = "", y = "Percent exports from unknown source") +
  facet_wrap(~iso3c) +
  theme_bw()

supply %>%
  group_by(year, iso3c) %>%
  summarise(error_export = sum(error_export, na.rm = TRUE), 
            domestic_export = sum(domestic_export, na.rm = TRUE),
            foreign_export = sum(foreign_export, na.rm = TRUE),
            error_percent = 100*sum(error_export, na.rm = TRUE)/(sum(domestic_export, na.rm = TRUE) + 
                                                     sum(foreign_export, na.rm = TRUE) + 
                                                     sum(error_export, na.rm = TRUE))) %>%
  group_by(iso3c) %>%
  mutate(total_error = sum(error_export)) %>%
  ungroup() %>%
  slice_max(n = 12*length(unique(supply$year)), order_by = total_error) %>%
  ggplot(aes(x = year, y = error_export)) +
  geom_line() +
  labs(x = "", y = "Exports from unknown source (t, live weight)") +
  facet_wrap(~iso3c) +
  theme_bw()

```

```{r explore_species_error, eval=FALSE}
# Plot time series of species with largest volume error exports
supply %>%
  group_by(year, sciname, CommonName) %>%
  summarise(error_export = sum(error_export, na.rm = TRUE), 
            domestic_export = sum(domestic_export, na.rm = TRUE),
            foreign_export = sum(foreign_export, na.rm = TRUE),
            error_percent = 100*sum(error_export, na.rm = TRUE)/(sum(domestic_export, na.rm = TRUE) + 
                                                     sum(foreign_export, na.rm = TRUE) + 
                                                     sum(error_export, na.rm = TRUE))) %>%
  # Remove groupings only used by error 
  filter(!(sciname %in% c("animalia", "actinopteri"))) %>%
  group_by(sciname, CommonName) %>%
  mutate(total_error = sum(error_export)) %>%
  ungroup() %>%
  slice_max(n = 12*length(unique(supply$year)), order_by = total_error) %>%
  ggplot(aes(x = year, y = error_percent)) +
  geom_line() +
  labs(x = "", y = "Percent exports from unknown source") +
  facet_wrap(~CommonName) +
  theme_bw()

supply %>%
  group_by(year, sciname, CommonName) %>%
  summarise(error_export = sum(error_export, na.rm = TRUE), 
            domestic_export = sum(domestic_export, na.rm = TRUE),
            foreign_export = sum(foreign_export, na.rm = TRUE),
            error_percent = 100*sum(error_export, na.rm = TRUE)/(sum(domestic_export, na.rm = TRUE) + 
                                                     sum(foreign_export, na.rm = TRUE) + 
                                                     sum(error_export, na.rm = TRUE))) %>%
  # Remove groupings only used by error 
  filter(!(sciname %in% c("animalia", "actinopteri"))) %>%
  group_by(sciname, CommonName) %>%
  mutate(total_error = sum(error_export)) %>%
  ungroup() %>%
  slice_max(n = 12*length(unique(supply$year)), order_by = total_error) %>%
  ggplot(aes(x = year, y = error_export)) +
  geom_line() +
  labs(x = "", y = "Exports from unknown source (t, live weight)") +
  facet_wrap(~CommonName) +
  theme_bw()

```

# Processing hotspots

```{r}
artis %>% 
  filter(dom_source == "foreign export",
         year %in% 2015:2019) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)/5) %>%
  mutate(country_name = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  mutate(country_name = fct_reorder(country_name, live_weight_t)) %>%
  slice_max(n = 25, order_by = live_weight_t) %>%
  ggplot(aes(x = live_weight_t, y = country_name)) +
  geom_bar(stat = "identity") +
  labs(title = "Average exports of foreign origin (2015-2019)",
       x = "Foreign exports (t)", y = "") +
  theme_bw()

artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(exporter_iso3c) %>%
  mutate(total_export = sum(live_weight_t)) %>%
  ungroup() %>%
  filter(dom_source == "foreign export") %>%
  group_by(exporter_iso3c) %>%
  summarise(export_percent_foreign = 100*sum(live_weight_t)/mean(total_export)) %>%
  mutate(country_name = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  mutate(country_name = fct_reorder(country_name, export_percent_foreign)) %>%
  slice_max(n = 25, order_by = export_percent_foreign) %>%
  ggplot(aes(x = export_percent_foreign, y = country_name)) +
  geom_bar(stat = "identity") +
  labs(title = "Average percent exports of foreign origin (2015-2019)",
       x = "Foreign exports as percent of total exports", y = "") +
  theme_bw()

artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(importer_iso3c) %>%
  mutate(total_export = sum(live_weight_t)) %>%
  ungroup() %>%
  filter(dom_source == "foreign export") %>%
  group_by(exporter_iso3c) %>%
  summarise(export_percent_foreign = 100*sum(live_weight_t)/mean(total_export)) %>%
  mutate(country_name = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  mutate(country_name = fct_reorder(country_name, export_percent_foreign)) %>%
  slice_max(n = 25, order_by = export_percent_foreign) %>%
  ggplot(aes(x = export_percent_foreign, y = country_name)) +
  geom_bar(stat = "identity") +
  labs(title = "Average percent of imports re-exported (2015-2019)",
       x = "Foreign exports as percent of total imports", y = "") +
  theme_bw()

```