### 2023-01-20 Archive/ms_figures.Rmd
archived results and figures

---
title: "ms_figures"
author: "Jessica Gephart"
date: "2022-08-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(countrycode)
library(circlize)
library(here)
library(ggpubr)
library(kableExtra)
library(DBI)
library(janitor)
library(viridis)
library(broom)
library(exploreARTIS)
library(vegan)
library(tidytext)
```

```{r load_database_data}
# Load ARTIS data from SQL local database
con <- dbConnect(RPostgres::Postgres(),
                 dbname=Sys.getenv("POSTGRES_DB"),
                 host="localhost",
                 port="5432",
                 user=Sys.getenv("POSTGRES_USER"),
                 password=Sys.getenv("POSTGRES_PASSWORD"))

# Check that connection is established by checking which tables are present
dbListTables(con)

# Pull all ARTIS and production data
artis <- dbGetQuery(con, "SELECT * FROM snet")

artis <- artis %>%
  select(-record_id)

prod <- dbGetQuery(con, "SELECT * FROM production")
prod <- prod %>%
  select(-record_id)

country_metadata <- dbGetQuery(con, "SELECT * FROM countries")
country_metadata <- country_metadata %>%
  select(-record_id)

code_max_resolved_taxa <- dbGetQuery(con, "SELECT * FROM code_max_resolved_taxa") %>%
              mutate(hs6 = as.character(hs6)) %>%
  select(-record_id)

sciname_metadata <- dbGetQuery(con, "SELECT * FROM sciname") %>%
  select(-record_id)

# Close database connection
dbDisconnect(con)

rm(list = c("con"))
```

```{r set_directories}
# Set directories 
outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization"
outdir <- "outputs"

# Set color palettes
# All colors from palette: #FFE6BE (light yellow), #F7AF75 (light orange), #F37542 (medium orange), 
# #E24027 (dark orange), #0F2D59 (navy), #1B708F (blue), #2E8D9A (teal), #86ADA7 (light blue)
method_colors <- c("#F7AF75", "#2E8D9A", "#86ADA7") # light orange, teal, light blue
source_colors <- c("#F37542", "#0F2D59", "#86ADA7") # medium orange, navy, light blue
environment_colors <- c( "#2E8D9A", "#0F2D59","#86ADA7") # teal, navy, light blue

habitat_source_colors <- c("#1B708F", # blue = marine capture
                          "#86ADA7", # light blue = freshwater capture
                          "#F37542", # medium orange = marine aquaculture
                          "#F7AF75", # light orange = inland aquaculture
                          "grey50" # unknown
                          )

region_colors <- c(
  "#741A32", # maroon
  "#B34232", # burnt orange
  "#D38F35", #  orange
  "#D4B95f", # khaki
  "#4FA2A2", # teal
  "#114F59" # dark teal
)

```

```{r load_drive_data}

pop <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/fao_annual_pop.csv") %>% 
  left_join(country_metadata, by = c("iso3c")) %>%
  select(iso3c, year, "region" = "owid_region", pop)

```

```{r join_data}
# Summarize cleaned FAO production data
prod <- prod %>% 
  select("year", "iso3c", "sciname", 
         "method", "habitat", 
         "production_t" = "live_weight_t") %>%
  group_by(year, iso3c, sciname, method, habitat) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE)) %>%
  ungroup()

# Join production data to ARTIS
artis <- artis %>%
  filter(live_weight_t >= 0.1) %>%
  # Add importer and exporter region
  left_join(country_metadata %>% 
              select("exporter_iso3c" = "iso3c", "exporter_region" = "owid_region"), by = c("exporter_iso3c")) %>%
  left_join(country_metadata %>% 
              select("importer_iso3c" = "iso3c", "importer_region" = "owid_region"), by = c("importer_iso3c")) %>%
# FIX IT: Determine if this should be deleted. Production of only species exports may not be relevant
  left_join(prod, by = c("year", "exporter_iso3c" = "iso3c", 
                           "sciname", "method", "habitat")) %>%
  ungroup()

# Update artis data to lowest resolution by comparing code taxa_level and prod taxa_level
artis <- artis %>%
  left_join(code_max_resolved_taxa,
            by = c("hs_version", "hs6", "sciname")) %>%
  left_join(sciname_metadata, 
            by = "sciname") %>%
  # Leave any missing sciname_hs_modified as original sciname
  mutate(sciname_hs_modified = case_when(
    is.na(sciname_hs_modified) ~ sciname,
    TRUE ~ sciname_hs_modified
  )) %>%
  rename(environment = habitat) %>%
  # Add habitat-method column
  mutate(habitat_method = paste(environment, method, sep =" ")) %>% 
  mutate(habitat_method = case_when(
    str_detect(habitat_method, "unknown") ~ "unknown", 
    TRUE ~ habitat_method
  )) %>% 
  # Set factor levels
  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
                                                            "marine aquaculture", "inland aquaculture",
                                                            "unknown" )))

# Remove data frames after merged with ARTIS
rm(list = c("code_max_resolved_taxa", "sciname_metadata"))

```

# Figure 1: Increases in global export of blue foods

## Figure 1 a: Exports of global marine and freshwater aquaculture and fishery products (t live weight equivalent) from 1996-2019

```{r fig1a}
global_export_by_source <- artis %>% 
  mutate(food_or_fmfo = case_when(
    hs6 == "230120" ~ "fishmeal",
    hs6 != "230120" ~ "nonfishmeal"
  )) %>%
  group_by(year, habitat_method, food_or_fmfo) %>%
  summarise(live_weight_t = sum(live_weight_t))

write.csv(global_export_by_source, "outputs/global_export_by_source.csv", row.names = FALSE)

fig1a <- global_export_by_source %>%
  group_by(year, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
  geom_area() +
  scale_fill_manual(values = habitat_source_colors) +
  labs(y = "Exports (mil. t, live weight)", fill = "Source", x = "") +
  theme_bw() +
  theme(legend.position = "bottom")

```

## Figure 1 b: Percent of global marine and freshwater aquaculture and fishery production exported, excluding re-exports

```{r fig1b}
prod_by_source <- prod %>%
  group_by(year, method, habitat) %>% 
  summarise(production_t = sum(production_t)) %>%
  mutate(habitat_method = paste(habitat, method, sep = " ")) %>%
    # Set factor levels
  mutate(habitat_method = factor(habitat_method, 
                                 levels = c("marine capture", "inland capture",
                                            "marine aquaculture", "inland aquaculture"))) %>%
  ungroup() %>%
  select(-habitat, -method)

dom_export_by_source <- artis %>% 
  filter(dom_source == "domestic export") %>%
  group_by(year, habitat_method, environment) %>%
  summarise(dom_live_weight_t = sum(live_weight_t))

percent_export_by_source <- dom_export_by_source %>%
  left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
  mutate(percent_export = 100*dom_live_weight_t/production_t)

write.csv(percent_export_by_source, "outputs/percent_export_by_source_summary.csv", row.names = FALSE)

percent_export_total <- dom_export_by_source %>%
  left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
  group_by(year) %>% 
  summarise(dom_live_weight_t = sum(dom_live_weight_t), 
            production_t = sum(production_t)) %>%
  mutate(percent_export = 100*dom_live_weight_t/production_t) %>% 
  mutate(calc_type = "domestic exports") %>% 
  select(year, percent_export, calc_type)

percent_export_total_all_source <- artis %>% 
  group_by(year) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  left_join(prod_by_source %>% 
              group_by(year) %>% 
              summarise(production_t = sum(production_t)), 
            by = c("year")) %>% 
  mutate(percent_export = 100*live_weight_t/production_t) %>% 
  mutate(calc_type = "all export sources") %>% 
  select(year, percent_export, calc_type)

percent_export_total <- percent_export_total %>%
  bind_rows(percent_export_total_all_source)

write.csv(percent_export_total, "outputs/percent_export_total_summary.csv", row.names = FALSE)

fig1b <- ggplot() +
  geom_line(data = percent_export_by_source, aes(x = year, y = percent_export, color = habitat_method)) +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "% production exported", x = "", color = "Source") +
  theme_bw()

```

## Figure 1 c: Degree or measure of degree distribution by production method and environment

```{r fig1c}
all_degree_summary <-  artis %>% 
  select(year, exporter_iso3c, importer_iso3c) %>%
  distinct() %>%
  group_by(year, exporter_iso3c) %>%
  tally() %>% 
  rename(all_out_degree = n, iso3c = exporter_iso3c) %>%
  full_join(artis %>% 
              select(year, exporter_iso3c, importer_iso3c) %>%
              distinct() %>%
              group_by(year, importer_iso3c) %>%
              tally() %>%
              rename(all_in_degree = n, iso3c = importer_iso3c), 
            by = c("year", "iso3c"))

write.csv(all_degree_summary, "outputs/all_degree_summary.csv", row.names = FALSE)

habitat_method_degree_summary <-  artis %>% 
  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
  distinct() %>%
  group_by(year, exporter_iso3c, habitat_method) %>%
  tally() %>% 
  rename(all_out_degree = n, iso3c = exporter_iso3c) %>%
  full_join(artis %>% 
              select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
              distinct() %>%
              group_by(year, importer_iso3c, habitat_method) %>%
              tally() %>%
              rename(all_in_degree = n, iso3c = importer_iso3c), 
            by = c("year", "iso3c", "habitat_method"))

write.csv(habitat_method_degree_summary, "outputs/habitat_method_degree_summary.csv", row.names = FALSE)

fig1c_out_degree <- artis %>% 
  filter(habitat_method != "unknown") %>%
  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
  distinct() %>%
  group_by(year, exporter_iso3c, habitat_method) %>%
  tally() %>%
  group_by(year, habitat_method) %>%
  ggplot(aes(x = year, y = n, color = habitat_method)) +
  geom_smooth() +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "Out degree", fill = "Source") + 
  theme_bw()

fig1c_in_degree <- artis %>% 
  filter(habitat_method != "unknown") %>%
  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
  distinct() %>%
  group_by(year, importer_iso3c, habitat_method) %>%
  tally() %>%
  group_by(year, habitat_method) %>%
  ggplot(aes(x = year, y = n, color = habitat_method)) +
  geom_smooth() +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "In degree", fill = "Source") + 
  theme_bw()

```

```{r fig1, fig.height = 3, fig.width = 4.5}
png("outputs/fig1.png")
ggarrange(fig1a,
          ggarrange(fig1b,
                    fig1c_out_degree,
                    ncol = 2, nrow = 1,
                    legend = "none",
                    labels = c("b", "c")),
          common.legend = TRUE, legend = "bottom",
          ncol = 1, nrow = 2, labels = c("a", ""))

dev.off()
```

# Figure 2: Trade patterns

```{r SI_top_capture_aqua_imp_exp, fig.height=4, fig.width=4}
# Summarize total imports and exports by habitat and environment
bilateral_habitat_method_summary <- artis %>%
  # Remove if we decide to leave FMFO in
  filter(hs6 != "230120") %>%
  group_by(year, exporter_iso3c, exporter_region, importer_iso3c, importer_region, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t))

write.csv(bilateral_habitat_method_summary, "outputs/bilateral_habitat_method_summary.csv", row.names = FALSE)

top_traders_recent <- bilateral_habitat_method_summary %>% 
  filter(year %in% 2016:2020, habitat_method != "unknown") %>%
  group_by(importer_iso3c, exporter_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2016:2020))

top_importers_recent <- top_traders_recent %>%
  group_by(importer_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = live_weight_t) %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  mutate(ranking = rank(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = reorder_within(importer_iso3c, ranking, habitat_method)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

top_exporters_recent <- top_traders_recent %>%
  group_by(exporter_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = live_weight_t) %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  mutate(ranking = rank(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = reorder_within(exporter_iso3c, ranking, habitat_method)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_exporters_recent, top_importers_recent, ncol = 2, labels = c("Export", "Import"))

top_traders_past <- bilateral_habitat_method_summary %>% 
  filter(year %in% 1996:2000, habitat_method != "unknown") %>%
  group_by(importer_iso3c, exporter_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000))

top_importers_past <- top_traders_past %>%
  group_by(importer_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = live_weight_t) %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  mutate(ranking = rank(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = reorder_within(importer_iso3c, ranking, habitat_method)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

top_exporters_past <- top_traders_past %>%
  group_by(exporter_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = live_weight_t) %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  mutate(ranking = rank(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = reorder_within(exporter_iso3c, ranking, habitat_method)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_exporters_past, top_importers_past, ncol = 2, labels = c("Export", "Import"))

# Largest export increases
top_exporter_change <- bilateral_habitat_method_summary %>% 
  filter(year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_iso3c, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = end-beginning) %>% 
  arrange(desc(change))

# Largest increases
top_export_increases <- top_exporter_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(exporter_iso3c = reorder_within(exporter_iso3c, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

# Largest decreases
top_export_decreases <- top_exporter_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_min(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(exporter_iso3c = reorder_within(exporter_iso3c, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_export_decreases, top_export_increases, nrow = 1)

# Largest import increases
top_importer_change <- bilateral_habitat_method_summary %>% 
  filter(year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, importer_iso3c, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = end-beginning) %>% 
  arrange(desc(change))

# Largest increases
top_import_increases <-top_importer_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(importer_iso3c = reorder_within(importer_iso3c, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

# Largest decreases
top_import_decreases <-top_importer_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_min(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(importer_iso3c = reorder_within(importer_iso3c, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_import_decreases, top_import_increases, nrow = 1)

# Repeat for regional trends
# Largest export increases
top_exporter_change <- bilateral_habitat_method_summary %>% 
  filter(year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_region, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = end-beginning) %>% 
  arrange(desc(change))

# Largest increases
top_region_export <- top_exporter_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(exporter_region = reorder_within(exporter_region, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = exporter_region, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=exporter_region, yend=exporter_region)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

# Largest import increases
top_importer_change <- bilateral_habitat_method_summary %>% 
  filter(year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, importer_region, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = end-beginning) %>% 
  arrange(desc(change))

# Largest increases
top_region_import <-top_importer_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(importer_region = reorder_within(importer_region, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = importer_region, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=importer_region, yend=importer_region)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_region_export, top_region_import, nrow = 1, labels = c("Export", "Import"))

```

```{r SI_top_traders}
total_country_exports <- artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(exporter_iso3c, method, environment) %>%
  summarise(export_live_weight_t = sum(live_weight_t)/length(2015:2019)) %>%
  mutate(export_live_weight_t = export_live_weight_t/1000000) %>%
  rename("iso3c" = "exporter_iso3c")

total_country_imports <- artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(importer_iso3c, method, environment) %>%
  summarise(import_live_weight_t = sum(live_weight_t)/length(2015:2019)) %>%
  mutate(import_live_weight_t = import_live_weight_t/1000000) %>%
  rename("iso3c" = "importer_iso3c")

total_country_trade <- total_country_exports %>%
  full_join(total_country_imports, by = c("iso3c", "method", "environment")) %>%
  mutate(trade_live_weight_t = export_live_weight_t + import_live_weight_t) %>%
  ungroup() %>%
  group_by(iso3c) %>%
  mutate(total = sum(trade_live_weight_t)) %>%
  ungroup() %>%
  slice_max(n = 9*20, order_by = total)

total_net_import <- total_country_trade %>%
  group_by(iso3c) %>%
  summarise(export_live_weight_t = sum(export_live_weight_t),
            import_live_weight_t = sum(import_live_weight_t)) %>%
  mutate(net_import = import_live_weight_t - export_live_weight_t,) %>%
  select(iso3c, net_import)

total_country_trade %>% 
  left_join(total_net_import, by = "iso3c") %>%
  mutate(iso3c = fct_reorder(as.factor(iso3c), -net_import)) %>%
  ggplot() +
  geom_bar(aes(y = -1*export_live_weight_t, x = iso3c, fill = method), stat = "identity", position = "stack") +
  geom_bar(aes(y = import_live_weight_t, x = iso3c, fill = method), stat = "identity", position = "stack") +
  scale_fill_manual(values = method_colors) +
  geom_line(aes(y = net_import, x = iso3c), group = 1) +
  geom_hline(yintercept = 0) +
  labs(x = "", y = "Trade (million t, live weight)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

## Regional bilateral flows
```{r region_trend_matrix_by_method, fig.height=5, fig.width=5}
artis_region_method <- artis %>%
  filter(!is.na(importer_region), !is.na(exporter_region), 
         importer_region != "Other nei", exporter_region != "Other nei") %>%
  group_by(year, exporter_region, importer_region, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_importer = paste(exporter_region, " to ", importer_region, sep = ""))

# Create basic plot for the legend

# artis_region_method %>%
#   group_by(year, exporter_region) %>%
#   summarise(live_weight_t = sum(live_weight_t)) %>%
#   ggplot(aes(x = year, y = live_weight_t, fill = exporter_region)) +
#   geom_area() +
#   scale_fill_manual(values = region_colors) +
#   labs(fill = "Region") +
#   theme(legend.position = "bottom")
```

```{r regional_bilateral_trends}
artis_region <- artis_region_method %>%
  group_by(year, exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_importer = paste(exporter_region, " to ", importer_region, sep = ""))

region_slopes <- artis_region %>%
  split(.$exporter_importer) %>% 
  map(~lm(live_weight_t~year, data = .x)) %>% 
  map_df(broom::tidy, .id = 'exporter_importer') %>%
  filter(term == 'year') %>%
  select(exporter_importer, "slope" = "estimate")

artis_region <- artis_region %>%
  left_join(region_slopes, by = "exporter_importer") %>% 
  mutate(intraregion = case_when(
           exporter_region == importer_region ~ "intraregional trade",
           exporter_region != importer_region ~ "interegional trade"))

```

# SI Fig X: Bilateral region trends lollypop plot

```{r region_trend_lollipop, fig.height=4, fig.width=3}
region_slopes_ranked <- artis_region %>%
  select(-year, -live_weight_t) %>%
  distinct() %>%
  mutate(exporter_importer = fct_reorder(exporter_importer, slope)) %>%
  ggplot(aes(x = slope/1000, y = exporter_importer, color = intraregion)) +
  geom_segment(aes(x=0, xend=slope/1000, y=exporter_importer, yend=exporter_importer, color = intraregion)) +
  geom_point(size=3) +
  scale_y_reordered() +
  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
  labs(y = "", x = "Ave. annual change in export (1000 t live weight)", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

region_slopes_ranked

```

```{r, fig.height=3, fig.width=5}
region_trend_matrix_method <- artis_region_method %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
  geom_area() +
  scale_fill_manual(values = habitat_source_colors) +
  labs(x = "", y = "Trade (million t, live weight)", color = "Ave. annual change in export \n(1000 t live weight)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        legend.position = "none",
        strip.background = element_blank(),
        plot.margin=unit(c(0,0,0,0), "cm")) +
  facet_grid(rows = vars(exporter_region), cols = vars(importer_region), 
             labeller = label_wrap_gen(width = 5), scales = "free")

exports_by_region <- artis_region %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = importer_region)) +
  geom_area() +
  labs(x = "", y = "", fill = "Importer") +
  scale_fill_manual(values = region_colors) +
  facet_wrap(~exporter_region, ncol = 1) +
  theme_bw() +
    theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.margin=unit(c(0,0.5,0,0), "cm"))

imports_by_region <- artis_region %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = exporter_region)) +
  geom_area() +
  labs(x = "", y = "", fill = "Exporter") +
  scale_fill_manual(values = region_colors) +
  facet_wrap(~importer_region, nrow = 1) +
  theme_bw() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.text.x = element_blank(),
        plot.margin=unit(c(0.5,0,0,0), "cm"))

png("outputs/fig2a.png", width = 6.5, height = 5, res = 500, units = "in")
ggarrange(ggarrange(imports_by_region, NULL,
                    ncol = 2, nrow = 1,
                    widths = c(4.5, 1.25), labels = c("a", "")),
          ggarrange(region_trend_matrix_method, 
                    ggarrange(NULL, exports_by_region,
                              ncol = 1, heights = c(0.3, 4)), 
                    ncol = 2, nrow = 1,
                    widths = c(4.5, 1), labels = c("b", "c")),
          ncol = 1, nrow = 2, 
          heights = c(1, 4.5))
dev.off()
```

# Figure 3: Supply patterns

```{r calculate_supply}
supply <- calculate_supply(artis %>% 
                             # Remove FM trade from supply calculations
                             #filter(hs6 != "230120") %>%
                             rename(habitat = environment), 
                           prod %>% 
                             rename(live_weight_t = production_t)) %>%
  # Add habitat-method column
  mutate(habitat_method = paste(habitat, method, sep =" ")) %>% 
  mutate(habitat_method = case_when(
    str_detect(habitat_method, "unknown") ~ "unknown", 
    TRUE ~ habitat_method
  )) %>% 
  # Set factor levels
  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
                                                            "marine aquaculture", "inland aquaculture",
                                                            "unknown" ))) %>% 
  left_join(country_metadata %>% 
              select(iso3c, "region" = "owid_region"), by = "iso3c") %>% 
      mutate(supply_no_error = case_when(supply_no_error < 0 ~ 0,
                                     TRUE ~ supply_no_error))

write.csv(supply, file.path(outdir, "supply.csv"), row.names = FALSE)
```

# Fig 3a: Foreign dependence trends

```{r supply_source_by_region}
region_supply_trends <- supply %>% 
  filter(region != "Other nei") %>%
  group_by(region, year) %>%
  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
  pivot_longer(cols = supply_domestic:supply_foreign, 
               names_to = "supply_source", values_to = "supply_percent") %>%
  mutate(supply_source = gsub("supply_", "", supply_source)) %>%
  ggplot(aes(x = year, y = supply_percent, color = supply_source)) +
  geom_line() +
  scale_color_manual(values = c("#F37542", "#0F2D59")) +
  labs(x = "", y = "Percent of supply", color = "Origin") +
  facet_wrap(~region) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

global_supply_trends <- supply %>% 
  filter(region != "Other nei") %>%
  group_by( year) %>%
  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
  pivot_longer(cols = supply_domestic:supply_foreign, 
               names_to = "supply_source", values_to = "supply_percent") %>%
  mutate(supply_source = gsub("supply_", "", supply_source)) %>%
  ggplot(aes(x = year, y = supply_percent, color = supply_source)) +
  geom_line() +
  scale_color_manual(values = c("#F37542", "#0F2D59")) +
  labs(x = "", y = "Percent of supply", color = "Origin") +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

supply %>% 
  filter(region != "Other nei") %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
  arrange(desc(supply_foreign))

```

```{r}
# Calculate per capita supply by region and source
supply_total <- supply %>% 
  group_by(iso3c, region, year, habitat_method) %>%
  summarise(supply = sum(supply),
            supply_no_error = sum(supply_no_error),
            supply_domestic = sum(supply_domestic),
            supply_foreign = sum(supply_foreign),
            import = sum(imports_live_weight_t), 
            domestic_export = sum(domestic_export),
            foreign_export = sum(foreign_export),
            error_export = sum(error_export), 
            production_t = sum(production_t)
            ) %>%
  left_join(pop %>% select(-region), by = c("iso3c", "year"))

# Global per capita supply
global_percap_supply_area <- supply_total %>%
  filter(region != "Other nei") %>%
  group_by(year, habitat_method) %>%
  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
  ggplot(aes(x = year, y = supply_per_cap, fill = habitat_method)) +
  geom_area() + 
  labs(y = "Per capita supply (kg)", x = "", fill = "Source") +
  scale_fill_manual(values = habitat_source_colors) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

global_percap_supply_line <- supply_total %>%
  filter(region != "Other nei") %>%
  group_by(year, habitat_method) %>%
  summarise(supply_per_cap = 1000 * (sum(supply_no_error))/sum(pop)) %>%
  ggplot(aes(x = year, y = supply_per_cap, color = habitat_method)) +
  geom_line() + 
  labs(y = "Per capita supply (kg)", x = "", color = "Source") +
  scale_color_manual(values = habitat_source_colors) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Regional per capita supply
regional_percap_supply_area <- supply_total %>%
  filter(region != "Other nei") %>%
  group_by(year, region, habitat_method) %>%
  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
  ggplot(aes(x = year, y = supply_per_cap, fill = habitat_method)) +
  geom_area() + 
  labs(y = "Per capita supply (kg)", x = "", fill = "Source") +
  scale_fill_manual(values = habitat_source_colors) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~region)

regional_percap_supply_line <- supply_total %>%
  filter(region != "Other nei") %>%
  group_by(year, region, habitat_method) %>%
  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
  ggplot(aes(x = year, y = supply_per_cap, color = habitat_method)) +
  geom_line() + 
  labs(y = "Per capita supply (kg)", x = "", color = "Source") +
  scale_color_manual(values = habitat_source_colors) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~region)

```

Where does trade have a positive versus negative effect on per capita supply for capture versus aquaculture?

```{r}
# Plot relationship between per capita supply and imports by prod source and region
country_supply <- supply %>% 
  filter(year == 2018, region != "Other nei") %>% 
  group_by(iso3c, region, year) %>% 
  summarise(supply = sum(supply),
            supply_no_error = sum(supply_no_error),
            supply_domestic = sum(supply_domestic),
            supply_foreign = sum(supply_foreign),
            import = sum(imports_live_weight_t), 
            domestic_export = sum(domestic_export),
            foreign_export = sum(foreign_export),
            error_export = sum(error_export), 
            production_t = sum(production_t)
  ) %>% 
  left_join(pop %>% select(-region), by = c("iso3c", "year")) %>%
  mutate(supply_per_cap = supply_no_error/pop) %>%
  # Converting from tonnes to kg
  mutate(supply_per_cap = 1000 * supply_per_cap)
```

```{r, fig.height = 5, fig.width = 3}
country_supply %>% 
  ungroup() %>%
  mutate(iso3c = fct_reorder(iso3c, supply_per_cap)) %>%
  #filter(supply_per_cap > 50) %>%
  ggplot(aes(x = supply_per_cap, y = iso3c, fill = region)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = region_colors) +
  theme_bw() +
  theme(legend.position = "none")

```

```{r}
import_supply <- country_supply %>%
  # FIX IT: Temporarily remove very large (unrealistic) supply values
  filter(supply_per_cap < 100) %>%
  ggplot(aes(x = import/1000, y = supply_per_cap, color = region)) +
  geom_point(alpha = 0.7) +
  scale_x_continuous(trans='log10') +
  labs(x = "Import (1000 t, live weight)", y = "Supply (per capita)", color = "Region") +
  scale_color_manual(values = region_colors) +
  theme_bw()

export_supply <- country_supply %>%
  # FIX IT: Temporarily remove very large (unrealistic) supply values
  filter(supply_per_cap < 100) %>%
  ggplot(aes(x = (domestic_export + foreign_export + error_export)/1000, y = supply_per_cap, color = region)) +
  geom_point(alpha = 0.7) +
  scale_x_continuous(trans='log10') +
  labs(x = "Export (1000 t, live weight)", y = "Supply (per capita)", color = "Region") +
  scale_color_manual(values = region_colors) +
  theme_bw()
```

```{r, fig.height = 5, fig.width = 7}
# Join figures
png("outputs/fig3.png", width = 7, height = 5, units = "in", res = 500)
ggarrange(
  ggarrange(global_percap_supply_line, regional_percap_supply_area, 
            common.legend = TRUE, nrow = 1, widths = c(1, 2),
            labels = c("a", "b")),
  ggarrange(global_supply_trends, region_supply_trends,
            common.legend = TRUE, nrow = 1, widths = c(1, 2),
            labels = c("c", "d")),
  nrow = 2
)
dev.off()
```

## Fig 4: Effect of trade on supply diversity

```{r supply_diversity}
# Calculate diversity indices
diversity <- supply %>%
  filter(supply_domestic > 0, supply_foreign > 0) %>%
  group_by(iso3c, region, year, sciname) %>%
  summarise(supply_no_error = sum(supply_no_error), 
            supply_domestic = sum(supply_domestic),
            supply_foreign = sum(supply_foreign),
            production_t = sum(production_t), 
            imports_live_weight_t = sum(imports_live_weight_t)) %>%
  ungroup(sciname) %>%
  mutate(shannon_supply_no_error = diversity(supply_no_error),
         shannon_supply_domestic = diversity(supply_domestic),
         shannon_supply_foreign = diversity(supply_foreign),
         shannon_prod = diversity(production_t),
         shannon_imports = diversity(imports_live_weight_t)) %>%
  select(iso3c, year, shannon_supply_no_error,
         shannon_supply_domestic, shannon_supply_foreign,
         shannon_prod, shannon_imports) %>%
  distinct()

total_trade <- supply %>%
  filter(supply_domestic > 0, supply_foreign > 0) %>%
  group_by(region, iso3c, year) %>%
  summarise(supply_domestic = sum(supply_domestic),
            supply_foreign = sum(supply_foreign),
            production_t = sum(production_t), 
            imports_live_weight_t = sum(imports_live_weight_t),
            domestic_export = sum(domestic_export),
            foreign_export = sum(foreign_export),
            error_export = sum(error_export))

diversity <- diversity %>%
  left_join(total_trade, by = c("region", "iso3c", "year"))

```

# Fig 4a: diversity of foreign and domestic supply

```{r}
# Compare diversity of foreign and domestic supply
foreign_vs_domestic_diversity <- diversity %>%
  filter(year == 2018, region != "Other nei") %>%
  ggplot(aes(x = shannon_supply_domestic, y = shannon_supply_foreign, color = region)) +
    geom_point(alpha = 0.7) +
    geom_abline(slope = 1, intercept = 0) +
  scale_color_manual(values = region_colors) +
  labs(x = "Domestic supply diversity", 
       y = "Foreign supply diversity",
       color = "") +
    theme_minimal()

# Slope less than 1 means domestic diversity is higher than foreign-sourced diversity
summary(lm(diversity$shannon_supply_foreign~diversity$shannon_supply_domestic))
```

# Fig 4b: Production and import diversity
```{r}
# Compare production and import diversity trends
prod_import_diversity_trends <- diversity %>%
  filter(region != "Other nei") %>%
  group_by(year, region) %>%
  summarise(shannon_prod = mean(shannon_prod), 
            shannon_imports = mean(shannon_imports)) %>%
  pivot_longer(cols = shannon_prod:shannon_imports,
               names_to = "source", values_to = "shannon") %>%
  mutate(source = case_when(
    source == "shannon_prod" ~ "Production diversity",
    source == "shannon_imports" ~ "Import diversity"
  )) %>%
  ggplot(aes(x = year, y = shannon, color = source)) +
  geom_line() +
  scale_color_manual(values = c(c("#F37542", "#0F2D59"))) +
  labs(x = "", y = "Shannon index", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~region)

# Compare total imports and supply diversity
imports_vs_supply_diversity <- diversity %>%
  filter(year == 2018, region != "Other nei") %>%
  ggplot(aes(x = imports_live_weight_t/1000000, y = shannon_supply_no_error, color = region)) +
  geom_point(alpha = 0.7) +
  scale_x_log10() +
  scale_color_manual(values = region_colors) +
  theme_minimal() +
  labs(x = "Total imports (million t, live weight)",
       y = "Supply diversity",
       color = "")

diversity %>%
  filter(region != "Other nei") %>%
  ggplot(aes(x = imports_live_weight_t/1000000, y = shannon_supply_no_error, color = region)) +
  geom_point(alpha = 0.5) +
  scale_x_log10() +
  scale_color_manual(values = region_colors) +
    labs(x = "Total imports (million t, live weight)",
       y = "Supply Shannon diversity",
       color = "") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  facet_wrap(~region)

diversity %>%
  mutate(dom_vs_foreign = case_when(
    shannon_supply_domestic >= shannon_supply_foreign ~ "dom",
    shannon_supply_domestic < shannon_supply_foreign ~ "foreign"
  )) %>% 
  group_by(dom_vs_foreign) %>%
  tally()

```

```{r, fig.height = 3, fig.width = 3}
png("outputs/fig4.png")
ggarrange(prod_import_diversity_trends, 
          ggarrange(foreign_vs_domestic_diversity,
                    imports_vs_supply_diversity,
                    nrow = 1,
                    common.legend = TRUE, 
                    legend = "bottom",
                    labels = c("b", "c")),
          nrow = 2,
          labels = c("a", "", "")
          )
dev.off()
```