### 2023-02-07 ms_results.Rmd
Update results with new text

---
title: "ms_results"
author: "Jessica Gephart"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
# Load packages
library(tidyverse)
library(countrycode)
library(kableExtra)
library(ggpubr)
library(tidytext)
```

```{r load_data}
# Directory
datadir <- "outputs_20230206"

global_export_by_source <- read.csv(file.path(datadir, "global_export_by_source.csv"))
all_degree_summary <- read.csv(file.path(datadir, "all_degree_summary.csv"))
habitat_method_degree_summary <- read.csv(file.path(datadir, "habitat_method_degree_summary.csv"))
```

# Trends blue food globalization

```{r trends_summary}
# Generate stats for text
total_1996 <- round(sum(global_export_by_source %>% filter(year == 1996) %>% pull(live_weight_t))/1000000, 1)
total_2019 <- round(sum(global_export_by_source %>% filter(year == 2019) %>% pull(live_weight_t))/1000000, 1)
total_2020 <- round(sum(global_export_by_source %>% filter(year == 2020) %>% pull(live_weight_t))/1000000, 1)

# Total trade by environment and production method
# Total trade for marine capture
total_marine_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "marine capture")) %>% pull(live_weight_t))/1000000, 1)
total_marine_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "marine capture")) %>% pull(live_weight_t))/1000000, 1)
total_inland_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "inland capture")) %>% pull(live_weight_t))/1000000, 1)
total_inland_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "inland capture")) %>% pull(live_weight_t))/1000000, 1)
# Total trade for marine aquaculture
total_marine_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_marine_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)
# Total trade for inland aquaculture
total_inland_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_inland_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)

# Total trade for all capture
total_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)
total_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)
total_capture_2020 <- round(sum(global_export_by_source %>% filter(year == 2020, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)

# Total trade for all aquaculture
total_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_aquaculture_2020 <- round(sum(global_export_by_source %>% filter(year == 2020, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)

# Total trade for FMFO
artis_fm <- read.csv(file.path(datadir, "artis_fm.csv"))

total_fmfo_1996 <- artis_fm %>%
  filter(year == 1996) %>%
  mutate(live_weight_t = live_weight_t/1000000) %>%
  pull(live_weight_t) %>%
  sum(na.rm = TRUE)

total_fmfo_2019 <- artis_fm %>%
  filter(year == 2019) %>%
  mutate(live_weight_t = live_weight_t/1000000) %>%
  pull(live_weight_t) %>%
  sum(na.rm = TRUE)

```

```{r percent_export_by_source_trends}
percent_export_by_source <- read.csv(file.path(datadir, "domestic_export_percent_production_by_habitat_method.csv"))
percent_export_total <- read.csv(file.path(datadir, "total_export_percent_prod.csv"))
percent_export_total_dom_1996 <- percent_export_total %>% 
  filter(year == 1996, calc_type == "domestic exports") %>% 
  pull(percent_export)
percent_export_total_dom_2019 <- percent_export_total %>% 
  filter(year == 2019, calc_type == "domestic exports") %>% 
  pull(percent_export)
percent_export_total_dom_1996 <- percent_export_total %>% 
  filter(year == 1996, calc_type == "domestic exports") %>% 
  pull(percent_export)
percent_export_total_1996 <- percent_export_total %>% 
  filter(year == 1996, calc_type == "all export sources") %>% 
  pull(percent_export)
percent_export_total_2019 <- percent_export_total %>% 
  filter(year == 2019, calc_type == "all export sources") %>% 
  pull(percent_export)
percent_marine_capture_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "marine capture") %>%
  pull(percent_export)
percent_marine_capture_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "marine capture") %>%
  pull(percent_export)
percent_marine_aqua_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "marine aquaculture") %>%
  pull(percent_export)
percent_marine_aqua_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "marine aquaculture") %>%
  pull(percent_export)
percent_inland_capture_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "inland capture") %>%
  pull(percent_export)
percent_inland_capture_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "inland capture") %>%
  pull(percent_export)
percent_inland_aqua_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "inland aquaculture") %>%
  pull(percent_export)
percent_inland_aqua_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "inland aquaculture") %>%
  pull(percent_export)
```

```{r trends_concentration}
# Export / Imports concentration
cumulative_percent_threshold <- 75
export_concentration_n_countries <- read.csv(file.path(datadir, "export_concentration_n_countries.csv"))

import_concentration_n_countries <- read.csv(file.path(datadir, "import_concentration_n_countries.csv"))
```

```{r degree_summary_stats}
all_degree_1996 <- all_degree_summary %>%
  filter(year == 1996) %>% 
  summarise(all_out_degree = mean(all_out_degree, na.rm = TRUE)) %>%
  pull(all_out_degree)
all_degree_2019 <- all_degree_summary %>%
  filter(year == 2019) %>% 
  summarise(all_out_degree = mean(all_out_degree, na.rm = TRUE)) %>%
  pull(all_out_degree)
```

## Text for MS

Globalization describes the degree of international connectedness, which can be characterized by increasing flows of input, intermediate and final products among countries. Aquatic food exports increased `r round(100*((total_2019-total_1996)/total_1996),0)`%  (from `r total_1996` to `r total_2019` mil t; Fig 1a)  from 1996-2019. Over that period, both farmed and wild exports increased, though aquaculture grew faster, more than tripling, whereas capture exports grew by `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`%. Corresponding with the start of the COVID-19 pandemic, global aquatic food exports declined `r -1*round(100*((total_2020-total_2019)/total_2019), 0)`% overall in 2020 relative to 2019, with a `r -1*round(100*((total_capture_2020-total_capture_2019)/total_capture_2019),0)`% decline in capture exports, but a `r round(100*((total_aquaculture_2020-total_aquaculture_2019)/total_aquaculture_2019),1)`% increase in aquaculture exports (Fig 1a). Despite aquaculture comprising half of aquatic food production, capture fishery products constitute `r 100*round(total_capture_2019/total_2019, 1)`% of exports.

Another measure of the degree of globalization is the share of production exported. Domestic exports increased to `r round(percent_export_total_dom_2019, 1)`% of production in 2019, up from `r round(percent_export_total_dom_2019, 1)`% in 1996. while total exports reached 33% of all production in 2019. For comparison, the share of cereal production exported grew from around 10% in the late 1990s to 17% in the 2020s (Clapp 2022). Increasing marine capture export volumes despite stagnating production resulted in marine capture products having the greatest share of production destined for export (`r round(percent_marine_capture_exported_2019, 1)`% of production) and experienced the largest increases in the share exported, and largest increases in the share exported (Fig 2b). Aquaculture production more than doubled from 1996 to 2019, but aquaculture exports grew even faster, leading to an increasing share exported (Fig 2b). Despite increases, inland aquaculture still had the lowest share of production destined for export in 2019 (domestic exports represented only `r round(percent_inland_aqua_exported_2019, 1)`% of production) (Fig 1b). This finding clarifies prior debates about the orientation of aquaculture, while export trends suggest a need to consider international markets when crafting nutrition-sensitive policies (Golden et al. 2017; Belton et al 2018; Gephart et al. 2019).

Globalization exposes countries to external shocks, while also serving as a buffer against local shocks. Recent work exploring trade characteristics associated with systemic risk to shocks suggests higher exposure when networks are highly connected and concentrated, and when countries are more dependent on imports (Gephart et al. 2016;  Marchand et al. 2016; d’Amour et al. 2016). From 1996-2019, aquatic food trade became more connected with the average number of export partners increasing `r round(100*((all_degree_2019-all_degree_1996)/all_degree_1996),0)`%  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`; Fig 4c). Marine capture networks are most highly connected, followed by marine aquaculture, with inland capture and aquaculture trade being the least connected.

Since 1996, aquatic food exports and imports have become moderately less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019. Compare this with crops where just 7 countries and the EU account for 90% of wheat exports and just four countries accounting >80% of maize exports (Clapp 2022). The declining concentration is driven by capture fishery exports, whereas both inland and marine aquaculture exports became somewhat more concentrated (Fig 2d). Aquaculture exports are generally concentrated among fewer countries than capture exports, corresponding to the high concentration of aquaculture production to a few key regions. Similarly, the concentration of trade for individual species tends to be much higher. The divergent trends in trade features and differences among aquatic food groups suggests differences in the degree and types of trade shock risks across the sector. Such differences were observed in responses to COVID-19 (Love et al. 2021). However, understanding risk to shocks across foods is a priority area for future work (Davis et al. 2021).

Though aquatic food production, distribution, and consumption remain highly uneven (Hicks et al. 2022), we found declinding import concentration, with `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 (Fig 2e). The more dispersed import patterns are likely associated with growing populations and expanding middle classes and urbanization, particularly in low- and middle-income countries, which are responsible for increasing aquatic food demand in many countries (Naylor et al. 2021; SOFIA 2022). However, the relationship between aquatic food demand and income varies across aquatic foods, with demand generally increasing with income for higher quality fish, but falling for lower quality fish (Naylor et al. 2021).

![Figure 1: Increases in global export of aquatic foods. a) Exports of global marine and freshwater aquaculture and fishery products (t live weight equivalent) from 1996-2019. b) Percent of global marine and freshwater aquaculture and fishery production exported, excluding re-exports. c) Average number of export partners (out degree) by production method and environment. d) Number of countries comprising 75% of global exports by production method and environment with the global total in the black line.  e) Number of countries comprising 75% of global imports by production method and environment with the global total in the black line.](outputs/fig1.png)

## Other summary stats
Global seafood trade increased from `r total_1996` million t (live weight) in 1996 to `r total_2019` million t (live weight) in 2019, representing a `r round(100*((total_2019-total_1996)/total_1996),0)`% increase over that period.

Global capture fishery trade increased from `r total_capture_1996` million t (live weight) in 1996 to `r total_capture_2019` million t (live weight) in 2019, representing a `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`% increase over that period. In 1996, capture fishery trade represented `r round(100*total_capture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_capture_2019/total_2019)`% of all seafood exports.

Global aquaculture fishery trade increased from `r total_aquaculture_1996` million t (live weight) in 1996 to `r total_aquaculture_2019` million t (live weight) in 2019, representing a `r round(100*((total_aquaculture_2019-total_aquaculture_1996)/total_aquaculture_1996),0)`% increase over that period. In 1996, aquaculture fishery trade represented `r round(100*total_aquaculture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_aquaculture_2019/total_2019)`% of all seafood exports.

Aquatic foods have become increasingly globalized, with the percent of production exported increasing by `r 100*round((percent_export_total_dom_2019-percent_export_total_dom_1996)/percent_export_total_dom_1996, 1)` and the volume and number of trade partnerships approximately doubling. However, trends differ across aquatic food groups, with marine capture remaining the most highly globalized group, but with marine aquaculture showing faster trade growth.

Marine-sourced products dominate both capture and aquaculture exports, comprising `r round(100*total_marine_capture_2019/(total_marine_capture_2019+total_inland_capture_2019), 1)`% and `r round(100*total_marine_aquaculture_2019/(total_marine_aquaculture_2019+total_inland_aquaculture_2019), 1)`%, respectively. Aquatic food trade outpaced fishmeal (for non-human consumption) trade, which remained relatively stable, declining `r abs(round(100*(total_fmfo_2019-total_fmfo_1996)/(total_fmfo_1996), 1))`% over the time period. Consequently, fishmeal exports as a share of the sum of aquatic food and fishmeal exports declined from `r abs(round(100*(total_fmfo_1996)/(total_1996+total_fmfo_1996), 1))`% in 1996 to only `r abs(round(100*(total_fmfo_2019)/(total_2019+total_fmfo_2019), 1))`% in 2019.

# Shifts in global flows of blue foods

```{r}
bilateral_flow_change <- read.csv(file.path(datadir, "bilateral_flow_change.csv"))

percent_flow_increases <- 100*(nrow(bilateral_flow_change %>% filter(change > 0)))/nrow(bilateral_flow_change)
```

## Text for MS

Given the geographic patchiness of wild fish and aquaculture production, trade helps meet aquatic food demand in countries. Aquatic food imports are especially important where per capita demand is rising, aquaculture is limited, and wild fishery catch is stagnant. Corresponding to the geographical variability, we find the top importers, exporters, and bilateral flows to vary by habitat and farmed versus wild source, underscoring the importance of disaggregating trade (Fig S1-3). For example, Asia and Europe, and to a lesser extent, North America dominated marine capture and aquaculture trade networks in recent years (Fig S1). Since 1996, marine trade shifted toward Asia and Europe, with large increases in marine capture imports to and exports from Africa (Fig 2b).

At the country level, although some countries rank among the top traders across production methods, such as China for exports and the United Stated for imports, many countries are only top traders for one production method. Inland aquatic food trade is dominated by aquaculture, with the highest exports from Vietnam and China and highest imports by the United States, Japan, and South Korea (Fig S2-3). In general, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig S1, Fig S6).

Intraregional trade tends to be higher than interregional trade for all products due to shorter transport distances, historical ties, patterns of aquatic food preferences and established regional trade agreements (Natale et al. 2015). We find that this pattern generally holds for aquatic food trade as intraregional trade is the highest or second highest for all regions other than Oceania and South America. Since 1996, trade increased or remained approximately stable between nearly all regional trade pairs, other than within North America. At the country level, trade increased between `r round(percent_flow_increases, 1)`% of trade pairs. Despite trade increasing with partners across the globe, trade within Asia, Europe and Africa outpaced interregional trade with those regions. The largest average annual growth increases occurred for trade within Asia and within Europe, followed by trade between Europe and Asia (Fig 2a; Fig S6). Our trade estimates are ultimately from reported trade and therefore do not capture informal and unreported trade networks. Though estimated unreported trade is not globally available, it can be significant, especially for neighboring countries. For example, informal exports from Benin to Nigeria are estimated to be more than five times the formal exports (Benassi et al. 2019). Including informal trade would therefore likely strengthen intraregional trade patterns.

Across regions, Europe and North America have the highest net imports while South America has the highest net exports (Fig 2a). Least developed countries collectively are net exporters of aquatic foods across all production methods, with net exports almost tripling between 1996 and their 2018 peak (Fig S7). Least developed country net exports are dominated by marine capture products, which nearly quadrupled from 1996-2019. Net exports of aquatic foods may be economically beneficial to least developed countries where high value species are exported and revenue used to purchase other foods (Asche et al, 2015). However, economic and political barriers inhibit these wealth-based benefits from being realized (Hicks et al. 2022). Further, recent work exploring movement of nutrients derived from fisheries suggests international trade is driving redistribution of essential micronutrients prevalent and bioavailable in aquatic foods from areas of high deficiency in middle and low income countries to developed nations with greater nutrient security (Nash et al, 2022). Further, Nash et al (2022) found evidence of high vulnerability to future shifts in trade practices among middle and low income nations currently benefiting from trade in aquatic foods.

![Figure 2: Regional trade flows by production source (habitat and method). a) Total imports (positive) and exports (negative) colored by source, with net import trend in black. b) Bilateral flows colored by production source with exporting region along the rows and importing region along the columns. Values represent million tonnes in live weight equivalents. Note the scale difference across rows.](outputs/fig2.png)

## Other summary stats

# National blue food supplies
```{r}
supply <- read.csv(file.path(datadir, "supply.csv"))

global_supply <- read.csv(file.path(datadir, "global_supply.csv"))

global_supply_per_cap <- read.csv(file.path(datadir, "global_supply_per_cap.csv"))

regional_supply_per_cap <- read.csv(file.path(datadir, "fig3b_data.csv"))

# Pull global supply numbers from 1996 and 2019
global_supply_1996 <- global_supply_per_cap %>%
  filter(year == 1996) %>%
  pull(per_cap)

global_supply_2019 <- global_supply_per_cap %>%
  filter(year == 2019) %>%
  pull(per_cap)

# Check trends across regions
regional_supply_trends_summary <- regional_supply_per_cap %>%
  select(year, region, habitat_method, supply_per_cap) %>%
  group_by(year, region) %>% 
  summarise(supply_per_cap = sum(supply_per_cap, na.rm = TRUE)) %>%
  filter(year %in% c(1996, 2019)) %>%
  pivot_wider(names_from = year, values_from = supply_per_cap, names_prefix = "year") %>%
  mutate(diff = year2019-year1996, 
         percent_diff = 100*diff/year1996)

regional_supply_trends_habitatmethod_summary <- regional_supply_per_cap %>%
  select(year, region, habitat_method, supply_per_cap) %>%
  filter(year %in% c(1996, 2019), habitat_method != "unknown") %>%
  pivot_wider(names_from = year, values_from = supply_per_cap, names_prefix = "year") %>%
  mutate(diff = year2019-year1996, 
         percent_diff = 100*diff/year1996)

habitatmethod_supply_trends_summary <- regional_supply_per_cap %>%
  select(year, habitat_method, supply_per_cap) %>%
  group_by(year, habitat_method) %>% 
  summarise(supply_per_cap = sum(supply_per_cap, na.rm = TRUE)) %>%
  filter(year %in% c(1996, 2019), habitat_method != "unknown") %>%
  pivot_wider(names_from = year, values_from = supply_per_cap, names_prefix = "year") %>%
  mutate(diff = year2019-year1996, 
         percent_diff = 100*diff/year1996)

habitat_method_supply <- read.csv(file.path(datadir, "habitat_method_supply.csv"))

tmp <- habitat_method_supply %>%
  select(habitat, method, year, supply) %>%
  filter(year == 1996 | year == 2020) %>%
  mutate(year = paste("year", year, sep = "_")) %>%
  pivot_wider(names_from = year, values_from = supply) %>%
  mutate(percent_diff = 100 * (year_2020 - year_1996) / year_1996)

inland_aqua_supply_change <- tmp %>%
  filter(habitat == "inland" & method == "aquaculture") %>%
  pull(percent_diff)

marine_aqua_supply_change <- tmp %>%
  filter(habitat == "marine" & method == "aquaculture") %>%
  pull(percent_diff)

marine_cap_supply_change <- tmp %>%
  filter(habitat == "marine" & method == "capture") %>%
  mutate(percent_diff = abs(percent_diff)) %>%
  pull(percent_diff)

marine_cap_supply_percent <- habitat_method_supply %>%
  filter(year == 2020) %>%
  filter(percent_supply == max(percent_supply)) %>%
  pull(percent_supply)

regional_method_supplies <- read.csv(file.path(datadir, "regional_method_supplies.csv"))

oceania_cap_supply_percent <- regional_method_supplies %>%
  filter(region == "Oceania") %>%
  pull(capture)

asia_cap_supply_percent <- regional_method_supplies %>%
  filter(region == "Asia") %>%
  pull(capture)

method_supplies <- read.csv(file.path(datadir, "method_supplies.csv"))

cap_supply_percent <- method_supplies %>%
  filter(method == "capture") %>%
  pull(percent_supply)

consumption_foreign <- read.csv(file.path(datadir, "consumption_foreign.csv"))

foreign_consumption_start <- consumption_foreign %>%
  filter(year == 1996) %>%
  pull(percent_supply)

foreign_consumption_end <- consumption_foreign %>%
  filter(year == 2020) %>%
  pull(percent_supply)

asia_farmed_year <- supply %>%
  filter(region == "Asia") %>%
  group_by(year, method) %>%
  summarize(supply = sum(supply)) %>%
  ungroup() %>%
  filter(method == "capture" | method == "aquaculture") %>%
  pivot_wider(names_from = method, values_from = supply) %>%
  filter(aquaculture > capture) %>%
  filter(year == min(year)) %>%
  pull(year)

foreign_consumption_region <- read.csv(file.path(datadir, "fig3d_data.csv"))

```

```{r}
national_supply_diversity <- read.csv(file.path(datadir, "fig4b_data.csv"))

n_import_exceeds_prod <- national_supply_diversity %>%
  filter(shannon_imports > shannon_prod) %>%
  nrow()

```

## Text for MS

Since direct measurements of human consumption (e.g. dietary intake) are not collected globally, consumption is generally represented by apparent consumption. Apparent consumption is generally calculated as production plus imports minus exports and waste. Trade is therefore central to estimating consumption and historically limited understanding of aquatic food consumption patterns. By estimating species level trade, we are able to estimate apparent consumption of aquatic foods by species, production method, and geographical origin.

Globally, annual aquatic food apparent consumption increased from `r signif(global_supply_1996, digit = 3)` kg per capita in 1996 to `r signif(global_supply_2019, digit = 3)` kg per capita in 2019. Our estimates are slightly lower than FAOSTAT, which reports global aquatic food consumption at 15.6 kg/capita/year in 1996 and 20.7 kg/capita/year in 2019. We found that aquatic food consumption increased across all regions outside of North America, which was relatively stable and South America, where aquatic food consumption declined `r -1*signif(regional_supply_trends_summary %>% filter(region == "South America") %>% pull(percent_diff), digit = 3)`% (Fig 3b). Global increase were driven by inland and marine aquaculture, which increased by `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019) and `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019), respectively. Meanwhile, inland capture consumption grew modestly (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland capture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland capture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019, a `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(percent_diff), digits = 3)`% increase), while marine capture consumption declined `r -1*signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019). Nevertheless, capture still makes up `r signif(cap_supply_percent, digits = 2)`% of global aquatic food consumption, with its contribution to regional aquatic food consumption ranging from `r signif(oceania_cap_supply_percent, digits = 2)`% in Oceania to `r signif(asia_cap_supply_percent, digits = 2)`% in Asia, where farmed consumption overtook wild consumption in `r asia_farmed_year`.

Since we estimate exports from domestic and foreign sources, we can track changes in reliance on foreign sourced products. Globally the share of consumption that is foreign-sourced increased modestly, from `r signif(foreign_consumption_start, digits = 3)`% to `r signif(foreign_consumption_end, digits = 3)`% (Fig 3c). Patterns vary greatly across regions with Asia and S America dominated by domestic supply (`r signif(foreign_consumption_region %>% filter(year == 2019, region == "Asia", supply_source == "foreign") %>% pull(supply_percent), digits = 2)`% and `r signif(foreign_consumption_region %>% filter(year == 2019, region == "South America", supply_source == "foreign") %>% pull(supply_percent), digits = 2)`% foreign in 2019, respectively), but Europe dominated by foreign supply (`r signif(foreign_consumption_region %>% filter(year == 2019, region == "Europe", supply_source == "foreign") %>% pull(supply_percent), digits = 2)`% foreign in 2019) (Fig 3d). High reliance on foreign-sourced foods can pose a food security risk (Baer-Nawrocka and Sadowski 2019; Kummu et al. 2020), though it is not clear the extent to which these risks exist across aquatic foods. Nevertheless, countries have enacted policies to protect domestic supplies, including developing food stocks and subsidizing domestic food production (Wood et al. 2023). The United States used foreign dependence on aquatic foods as a motivation for a suite of policy changes to boost domestic production, including expanding aquaculture production and opening marine protected areas to fishing (Gephart et al. 2019).

High global diversity of aquatic food production enables high diversity in consumption, which is important for supplying a range of micronutrients, supporting a range of livelihoods, and is thought to support food system resilience (Seekell et al. 2017). However, one concern about food trade is that it homogenizes diets through two related pathways: by introducing foods that make diets more similar across countries and/or by out-competing diverse local foods. To explore this issue for aquatic foods, we compared the Shannon diversity index, of production versus imports and and exports (Fig 4a) and domestic supply versus foreign supply (Fig 4b). Across regions, import diversity is generally higher than production diversity and export diversity is similar to production diversity across regions (Fig 4a). Import diversity exceeds production diversity in `r n_import_exceeds_prod` out of `r nrow(national_supply_diversity)` countries. Although the diversity of imported goods tends to be higher than production, diversity of domestic-sourced consumption tends to be higher than  diversity of foreign sourced consumption (Fig 4b). Total imports and exports are positively associated with supply diversity across all regions (Fig 4c). This suggests trade is contributing new species to many national markets. Collectively, our diversity findings point to a need to protect domestic production diversity while still benefiting from the diversity of imported products.

## Other summary stats

![Figure 3: Aquatic food apparent consumption (supply) trends and regional patterns. a) Global aquatic food apparent consumption by production source over time. b) Regional aquatic food apparent consumption by production source over time. c) Global aquatic food domestic versus foreign sourcing over time. d) Regional aquatic food domestic versus foreign sourcing over time. ](outputs/fig3.png)

![Figure 4: Relationship between trade and supply diversity. a) Violin plots of the Shannon diversity of production, imports and exports for 2019. b) Violin plots of the Shannon diversity of domestic and foreign consumption for 2019. c) Relationship between total imports and overall supply diversity for 2019. d) Relationship between total exports and overall supply diversity for 2019. Note for figures 4c and 4d, only countries with a supply of at least 1 tonne were included.](outputs/fig4.png)